### 2017-04-05 paper/tab1.tex

\begin{table*}
\caption{Original Works Pertaining to Catalog Data\label{tab:sim_table}}
\begin{center}
\begin{tabular}{llll}
\hline
\hline
Original Reference & \# of Simulations & Code & Basic Physics \\
\hline
\citet{zuh10} & 4 & FLASH & N-body/Hydrodynamics \\
\citet{zuh11a} & 9 & FLASH & N-body/Hydrodynamics \\
\citet{zuh11b} & 6 & FLASH & Magnetohydrodynamics \\
\citet{zuh15} & 9 & Athena & Magnetohydrodynamics \\
\hline
\end{tabular}
\end{center}
\end{table*}